A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-09-11 reviewer_response.tex

\documentclass{aiaa-tc}

\title{A Graph Theoretic Approach to Problem Formulation for 
Multidisciplinary Design Analysis and Optimization \\  (Reviewer comments appear below in red text and the authors' responses appear in black text)}

@@

\begin{document}
\maketitle

\section{Reviewer 1}

\rev{This idea of PSG was mentioned but was not covered in 
the paper. For completeness and value proposition sake, I would recommend to 
add at least half section to show how it will help and look like. You can 
take the simplest formulation of your choice and show how MCG and FPG will 
translate and look as a PSG. A brief discussion of advantages vs 
disadvantages will be useful. }

We agree that the PSG was too quickly addressed in the paper. Although we feel that 
discussion of the PSG in significant depth would be too lengthy for the present paper, we agree that the topic does warrant a more complete discussion than that provided in the 
original draft.

We have added subsections 3.5 and 4.4  to address the the driver nodes and driven edges which 
together distinguish a PSG from an FPG.  We have also added subsection 5.5,
which provides an example of a simple PSG and compares and contrasts it with an XDSM diagram.

@@

\rev{ The idea of “holes” and “collision” is explained a few times in the paper. 
It can be reduced. }

In reviewing the paper, the authors agree that the concepts of holes and collisions are discussed perhaps too often.  The terms are formally defined only once 
in section 4.1. Section 5.3 provides a mathematical description of the set of nodes that meet
the criteria for holes and edges  in the context of the FPG algorithm, but it is our opinion that this discussion should stand separately from the definitions.

We have checked sections 4.1 and 5.3 and removed many instances of redundancy, including informal definitions repeated several times in the text.  Nonetheless, we feel that preserving the separate mathematical discussions in sections 4.1 and 5.3 makes for a clearer message in the paper. The concept is somewhat subtle 
since, in our definitions, a node can be either a design variable or a hole, but not both. Similarly, 
a collision can occur when a variable has two incomming edges if the problem is single fidelity but not if the problem 
is multi-fidelity. The additional discussion is intended to clarify theses subtleties.

@@

\rev{The review of other graph based techniques is good and highlights their 
limitations, but I would like to see a discussion of what challenges a 
practitioner may face while trying to use the approach suggested in this 
paper. The MCG and FPG will cover lot of information like objectives and 
constraints, but at what expense. Comparing XDSM with MCG or FPG, one might 
find XDSM as more easy to read and comprehend. For a similar size problem, 
FPG looks far more complex and bigger.}

We agree that XDSM is more visually informative and concise. However, an easily human-readable format was not the 
goal of the new syntax. Instead, the goal is to make a graph structure that 
was easier to operate on algorithmically to achieve feasible MDO data flows. As a consequence of this goal, the graph grows larger and becomes
less human readable. We believe that this tradeoff is acceptable for two reasons:

\begin{enumerate}
    \item It should be possible to convert between the gray lines of the XDSM format and the proposed syntax 
    in a lossless manner. 
    It is our conclusion that the PSG in our syntax represents a subset of the information in the XDSM diagram; namely, the data connectivity elements of XDSM and PSG are equivalent graphs.  The implication is that it would be possible to write an algorithm and associated code to "abstract" the FPG into a partial XDSM diagram that could then be elaborated by the user to add the solution-specific elements.  It may then be possible to write a PSG from an XDSM, if the additional information about all of the individual variable connections from the FPG graph were preserved in the background.
    (This is true for the gray lines of XDSM but not the black ones.
    Only data connections are shown in new syntax.) 
    \item Algorithmic methods are a useful way for analyzing MDO problem structure, so it seems appropriate to have a graph structure that works well for this use, even at the expense 
    of human readability. 
\end{enumerate}

Please see the revised end of the introduction (Section 1) in which the algorithmic purpose of the new graph syntax is now stated more clearly.

@@

\rev{If PSG is covered even to a small 
extent (like explained in first comment) I would suggest or like to see the 
PSG equivalent of problem formulation shown in Figure 2 with XDSM.}

We have added  Fig. 8 and Fig. 13 and associated discussion to address this request.

@@

\rev{Is it really adding a lot of value to have graphs like MCG or FPG which 
are very explicit in nature and can become difficult to read}

As discussed above, we believe that the value of the MCG and FPG is not derived from their ability to represent the data in a human readable 
form, but in their straightforward translation to machine readable formats. Section 5.3 introduces an 
algorithm for testing whether MDO data connectivity graphs are FPGs and for reducing graphs to FPGs. As problems grow larger and more complex, 
this type of algorithm is expected to become ever more valuable.

Such an algorithm could be implemented into a framework or stand-alone tool to help teams to develop the
specifics of a design problem before they begin detailed development work. 
Section 6 shows that for even simple problems, the graph (as viewed by an algorithm) is 
non-trivial. It also shows that there are multiple ways to solve problems. As 
problems grow even a small amount from the size of the example becomes invaluable. In sum, it is our belief that the graph syntax derives its usefulness from 
its effectiveness in describing MDO problem graphs in computational syntax.

\rev{Can the graph be compressed in a logical way to reduce the number of nodes 
and their interconnections while still conveying important and critical 
information for that stage (MCG stage or FPG stage) ? Reducing cycles is 
discussed but the graph may still look complex.
\\ \indent
Looking at these graphs (for example Fig. 8), a node for example Z2 
appears several times. It may be useful to figure out holes and collision, 
but once it is resolved, can those nodes be consolidated to provide simpler 
visualization ?}

There are a number of ways the graph could be simplified to make it more human readable. Variables could be combined, or they could be removed altogether such that the graph represents only component to component connections like a traditional DSM.

As discussed in the response above, it should be possible to convert the entire PSG graph into the data-connectivity part of an XDSM. This would present all of the information in a more compact and visually insightful form, e.g. an XDSM. By preserving the variable connectivity information, one could always convert back into the more verbose PSG syntax  if needed.  We would argue that the XDSM is more human readable, but the new syntax is a representation of how a framework tool, e.g. OpenMDAO or ModelCenter, could view the problem.

Given our goals, the verbose nature of the graphs is a necessary side effect.

@@

\rev{As the problem size increases, how will it affect the size and complexity 
of such graphs.}

The answer to this question is not trivial, as it depends on a number of factors such as the number of analyses and the number of variables in those analyses. However, roughly speaking, one could measure the complexity of a graph by the total number of nodes and edges. The number of nodes grows roughly linearly with the number of analysis tools and the number of variables. The number of fixed edges grows linearly with the number of variables as well. The number of connection edges depends entirely on the specifics of the problem, but, since any variable will have only a small finite number of incoming edges (usually 1, but maybe 2 or 3 in multi-fidelity case) one can avoid an unbounded growth. Therefore, very roughly speaking, the graph will grow as O(N+M) where N is the number of variables and M is the number of analyses.

If one assumes that large problems have hundreds of components and millions of variables, then graphs with a few million nodes and edges are possible. These numbers are very large, but are not outside of the realm of what is readily manageable and rapidly evaluated by today's graph libraries in common programming languages.

@@

\rev{
I will assume that creation of such a graph and interconnections will be 
expected to be done by some kind of software tool or wizard. A brief 
discussion about why this MCG, FPG and PSG idea is better to lead to such a 
system can be useful. Although this itself can be a separate research and 
paper, but why this idea can provide a better foundation can be valuable.
}

See the new section 2.4 for a short section describing our envisioned usage 
of the graph language. 
We have also mentioned in Section 5.3 that the algorithm can be directly implemented in any programming language capable of managing lists, such as Python.

@@

\rev{Minor comments:
Section 1 Line 46: I think other commercial tools such as iSight, 
modeFRONTIER should also be mentioned since their usage is more widespread 
compared to OpenMDAO.}

Agreed. We have now mentioned this tools alongside OpenMDAO in  Section 1.

\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “}

Fixed.

\rev{Page 16 line 38 
column 2: Typo : “… formulations involving MULTIPLES analysis ….”}

((let's discuss))
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\rev{Page 3, Line 21, 2nd column: change ``descrition'' to ``description''}

Fixed

\rev{Page 4, Line 37, 1st column: what does ``REMS'' stand for?}

``Reconfigurability in MDO Problem Synthesis''. The full name has been added to the paper.

@@

\rev{Page 4, Line 45, 1st column: change ``REMS address'' to ``REMS addresses''}

Fixed

\rev{Page 4, Lines 35 to 55, 1st column:  This paragraph is a bit awkward in 
structure.  In the second to last sentence you state, "REMS provides syntax 
for variables and functions but does not facilitate inclusion of solvers or 
optimizers in the graph."  The reader then expects your method to address 
these shortcomings in the next sentence but instead you state, "...the goal 
of allowing graphs to describe analysis couplings and more closely interact 
with existing design processes and MDAO frameworks," which you never really 
called out as a shortcoming of REMS in the first place.  I think you could 
improve the flow/wording here.}

The authors agree that this paragraph was poorly worded. We have re-worded it to attempt 
to make the issue with REMS more clear. The text now states (last two sentences, first column of p. 4): ``... REMS does not provide a mechanism for inclusion of solvers or optimizers in the graph. This fundamentally limits REMS from describing the specifics of different solution strategies as applied to a particular problem.''  We have removed the oddly juxtaposed statement about the goal of our graph syntax that was noted by the reviewer from the paragraph.  However, based on the comments by Reviewer 1 noted above, we have now added an additional description of how our method can capture optimizers and solvers in the context of the problem solution graph (PSG).

@@

\rev{Line 10, 1st column, Page 10:  This equation (missing a number) has a LOT of 
unions, variables, etc...in it.  It would help tremendously to provide a 
verbal description of at least one of these equations to help with 
translation into the real world.  This would be similar to what you did 
following the steps (1) to (4) in the next subsection.}

The authors agree.  We have added equation numbers (now Eqns. 8 and 9). We have also provided verbal descriptions to accompany the set notations.

@@

\rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the 
text here.  What criteria would be applied to determine which FPG is better 
than another?}

After considerable deliberation, we chose not to address the issue of how one should select one FPG in preference to another  in the context of the present paper. The answer 
to this question depends heavily on the specific situation. If one is limited by run-time of the MDO workflow, 
a smaller FPG with fewer cycles and variables may be preferable. If one is trying to resolve a detailed physical question, e.g. subtle design changes associated with nacelle-wing interference, then it may be preferable  to select a
more high fidelity and computationally expensive FPG. These notions are introduced in the example problem, but further development was beyond the scope of the paper.

The MCG provides a good foundation for finding the only FPG or indicate multiple possible FPGs. The user can then select the best FPG for the present stage of the design process, but this decision can  be revisited later in the process as the design goals change.

@@

\rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}

Fixed

\rev{Now for the request of more significant importance.  This is coming from 
someone with significant experience in MDO and who uses a lot of the 
analysis tools you specified in your case study on a daily basis.  I really 
liked this article but felt it was lacking some punch in the results 
section.  This was namely in the form of quantitative results.  You vaguely 
give some in the form of the different resulting connectivity graphs and 
this was good but to really drive home the value of your research to a 
potential user I think you could do a bit better with very minimal work. 
The potential user of this methodology will be concerned with one of two 
things: 1) minimizing computational time or 2) increasing the accuracy of 
the results.  I think you've given examples of these two scenarios already 
in figures 15 and 16, so why not provide a table of the CPU times (1) and/or 
deviation from higher order analyses (2) for each of these graphs.  Doesn't 
have to necessarily be these two metrics but something to really tie the 
output of your methodology to something tangible for comparison purposes. 
You could even include the metrics for a randomly obtained FPG as a 
baseline.  Then the person reading the article can say, "Wow, given a set of 
analysis tools and a general problem, I can employ this method to obtain an 
FPG that significantly reduces my overall computational time." or "Wow, 
given a set of analysis tools and a general problem, I can employ this 
method to obtain an FPG that significantly improves my overall computational 
accuracy while keeping my computational efficiency as high as possible."}

We agree that this sort of quantitative comparison would be very valuable. However, we 
believe that, in order not to be misleading, this comparison needs to be larger in scope than just that which could be obtained 
by executing the example problem given currently presented in the paper. We believe that one would need to address a wide range of problems with different graph structural 
characteristics and different inherent nonlinear features in the analyses in order to make statements that are more generalizable. The challenge is that quantitative measures are inherently 
dependent on particular details of a problem. 
For example, for coupled analyses, total CPU time is highly dependent on the number of iterations required to converge the system of nonlinear functions to a consistent solution within a specified tolerance.  This convergence is tied very closely to the specifics of the coupling and the nonlinearity.  We had previously considered the idea of selecting ``optimal FPGs'' in terms of aggregate metrics that could be specified \textit{a priori} as meta-data for a particular analysis, but defining these metrics without considering problem coupling and nonlinearity proved especially challenging.
To address this issue well would therefore necessitate the evaluation of many example problems with different characteristics.  This work, while very valuable, would 
likely require an entire paper in its own right. We thank the reviewer for this suggestion, and we consider it to be a great topic for future work in this area.

@@

\rev{If you are worried about length, I think the conclusions section may be cut 
in half and still accomplish its intended purpose.}

We agree that the conclusion section
could be more concise. We have shortened it by two paragraphs  and improved the wording.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The reviewer has correctly identified the major motivation for this work. The goal is indeed to 
assist engineers and teams of engineers in the process of building a model out of a set of interconnected 
tools. 
We believe that the distinction between phases 1 and 2 in the reviewer's  comments are covered adequately 
in the the introduction and especially in the notional problem described in the indented text on p. 2.
However, to clarify the distinction between phases 2 and 3 indicated by the reviewer, we modified the text following the indented text at the bottom of p. 2 to read as follows:

\emph{ ((copy and paste the last two paragraphs of Section 2.3))}

### 2013-09-11 conclusions.tex
Update with BG and DP edits to paper and reviewer response

The MCG addresses the question, ``Given a set of analysis tools, 
what are all of the variable interconnections between them that could be established?''  
The MCG provides a structured formalism to identify \textit{holes} and 
\textit{collisions} in interconnections between the set of analysis tools. 
In order to achieve a valid problem formulation, all of the holes and collisions must 
be resolved. We present an algorithm to perform this resolution which gurantees that you 
will find an FPG if one exists. In some cases this algorithm still relies on some input from the 
designers to make decisions. One benefit of a graph theoretic approach is the standard 
algorithms which can be used to inform the user's decisions, such as cycle detection, 
minimum spanning trees, and shortest path algorithms.

@@

For simple problems with few analysis tools and variables, the formulation of a 
valid problem description is straightforward. However, MDAO problems continue to 
grow in scale and complexity. As the numbers of analysis tools and variables have 
increased, it has become increasingly challenging and time consuming for engineering 
teams to determine how the multiple analysis tools can be interconnected to produce 
valid problem formulations, to know when other tools must be introduced, and to 
determine the number of free variables in the problem that can or should be 
varied by the designer or an optimizer. The graph formalism presented in this 
paper is intended to offer value in this context of increasing problem complexity by 
providing a formal approach to programatically identify and resolve missing and 
redundant data in order to create valid MDAO problem formulations.

>>>>

The MCG addresses the question, ``Given a set of analysis tools, 
what are all of the variable interconnections between them that could be established?''  
The MCG provides a structured formalism to identify \textit{holes} and 
\textit{collisions} in interconnections between the set of analysis tools. 
In order to achieve a valid problem formulation, all of the holes and collisions must 
be resolved. We present an algorithm to perform this resolution which guarantees that an FPG will be found if one exists. In some cases, this algorithm relies on additional input from the 
designers to make decisions. One benefit of a graph theoretic approach is the standard 
algorithms that can be used to inform the user's decisions, such as cycle detection, 
minimum spanning trees, and shortest path algorithms.

@@

For simple problems with few analysis tools and variables, the formulation of a 
valid problem description is straightforward. However, MDAO problems continue to 
grow in scale and complexity. As the numbers of analysis tools and variables have 
increased, it has become increasingly challenging and time consuming for engineering 
teams to determine how the multiple analysis tools can be interconnected to produce 
valid problem formulations, to know when other tools must be introduced, and to 
determine the number of free variables in the problem that can or should be 
varied by the designer or an optimizer. The graph formalism presented in this 
paper is intended to offer value in this context of increasing problem complexity by 
providing a formal approach to systematically identify and resolve missing and 
redundant data in order to create valid MDAO problem formulations.

### 2013-09-11 example_problem.tex
addresses second reviewers comments
Update with BG and DP edits to paper and reviewer response

	\item A variable node is created to represent the geometry variable as a input. Any inputs incorrectly omitted will be identified in step (D).

@@

	Although an FPG may now be obtained, this process has not provided a method to decide which edges to retain when resolving a collision, and these decisions will likely determine which analysis tools are included in the FPG. These decisions are left to the discretion of the designer in the context of the specific implementation. 
	However, the graph-based approach presented in this paper enables standard graph algorithms to be employed to automate these decisions based on considerations of metrics related to the graph or other data related to the analysis. 
	An example is a metric to obtain an FPG with the fewest possible number of cycles. 
In this example, the FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are inticated by dashed lines. 
For this example problem, the cycles were detected using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. 
There are two cycles which arise from each local output from PDCYL being directed into FLOPSa and then back into PDCYL.

>>>>

	\item A variable node is created to represent the geometry variable as an input. Any inputs incorrectly omitted will be identified in step (D).

@@

	Although an FPG may now be obtained, this process has not provided a method to decide which edges to retain when resolving a collision, and these decisions will likely determine which analysis tools are included in the FPG. These decisions are left to the discretion of the designer in the context of the specific implementation. 
	However, the graph-based approach presented in this paper enables standard graph algorithms to be employed to automate these decisions based on considerations of metrics related to the graph or other data related to the analyses. 
	An example is a metric to obtain an FPG with the fewest possible number of cycles. 
In this example, the FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are indicated by dashed lines. 
For this example problem, the cycles were detected using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. 
There are two cycles which arise from each local output from PDCYL being directed into FLOPSa and then back into PDCYL.
